---
title: 'Homework Five: Functions'
author: "April Wright"
date: "2022-09-23"
output: html_document
---
### Basanta Khakurel
## Due date: Sept 30, 5 PM.

```{r, eval=FALSE}
download.file(url = "https://raw.githubusercontent.com/BiologicalDataAnalysis2019/2022/main/vignettes/homework_5.Rmd", destfile = "C:/Users/basan/Desktop/Fall 2022/Biological_Data_Analysis/homeworks/homework_5.Rmd")
```
---

This homework concerns functions.

1. Make a function that reads in data and does some sort of manipulation of it. Undergrads, use the surveys dataset. Graduates, use your own data. Any manipulation is fine: selection, filtering, value removal. 

```{r}
select_data <- function(path_to_data){
  mydata <- read_csv(path_to_data)
  selected_data <- mydata %>%
    select(max_pairwise_difference, max_pairwise_difference_excluding_ambiguous, max_variable_block_length, max_variable_block_length_excluding_ambiguous, min_pairwise_difference, min_pairwise_difference_excluding_ambiguous, theta, tajima_d, tajima_pi, segregating_sites)
  return(View(selected_data))
}

```
***

2. Add a test to this function. How will you know if it worked, and alert a user if it did not?

```{r}
select_data <- function(path_to_data){
  mydata <- read_csv(path_to_data)
  selected_data <- mydata %>%
    select(max_pairwise_difference, max_pairwise_difference_excluding_ambiguous, max_variable_block_length, max_variable_block_length_excluding_ambiguous, min_pairwise_difference, min_pairwise_difference_excluding_ambiguous, theta, tajima_d, tajima_pi, segregating_sites)
  if (ncol(selected_data) == 10) {
    return(View(selected_data))
  } else {
    print("Please check your data file or path")
  }
}
```
***

3. Make a function that makes a plot. This can be any type of plot that we covered. Violin, box, scatter. Include the ability to save the plot to a file in the function.

```{r}
plot_func <- function(path_to_data){
  mydata <- read.csv(path_to_data)
  One <- ggplot(mydata, mapping = aes(x=1:nrow(mydata), y = max_pairwise_difference)) + geom_point()
  ggsave("Plots/Max_Pairwise_Difference.pdf", plot = One, dpi = 300, limitsize = FALSE)
  return(One)
}
```
***

4. For saving to a file, users of your function may want the ability to choose their file name. Make the filename an argument so that users can do so.

```{r}
plot_func <- function(path_to_data, filename.pdf){
  mydata <- read.csv(path_to_data)
  One <- ggplot(mydata, mapping = aes(x=1:nrow(mydata), y = max_pairwise_difference)) + geom_point()
  ggsave(filename.pdf, plot = One, dpi = 300, limitsize = FALSE)
  return(One)
}
```
***
***