---
title: "Homework One"
output: html_document
---

## Homework One

### Due Aug. 26th at 5pm. Simply write your answers in and save in the homeworks folder in your Rstudio.

This is going to be a low-stakes effort to get you comfortable with RStudio and RMarkdown. First, you're going to download the RMarkdown Document for this homework. Copy and paste this command into the console of your RStudio instance.

```{r}
download.file(url = "https://raw.githubusercontent.com/BiologicalDataAnalysis2019/2022/main/vignettes/homework_1.Rmd", destfile = "/cloud/project/homeworks/homework_1.Rmd")
```

Can you find where the file downloaded to? Look at the help entry for `download.file` if you aren't sure.

Open the file and answer these two questions: 

1. (1pt) What does `download.file` do? 

```
= It downloads the "homework_1.Rmd" file to save it locally.
```

2. (1pt) Where did your HW1.Rmd file download to? 

```
= The homework file is saved to /cloud/project/homeworks/.
```

Next, we're going to try to execute some code in an RMarkdown document. Press the green triangle.

```{r}
animals <- c("cat", "rat", "dog")  #this will save the values "cat", "rat", and "dog" as a vector to "animals"
weights <- c(15, 1, 120) #this will save the integers 15, 1 and 120 to the vector "weights".
animals[c(1,3)] # this will return us the first and the third values saved in the vector "animals".
```

What just happened? Annotate the above code with two comments (each worth one point) explaining the code.

We can also embed plots in RMarkdown. The below code should embed a bar plot. Not very useful, but let me know if you don't see one show up in your file.

```{r}
barplot(names.arg=animals, height=weights)
```

Did the plot show up OK?

```
Yes 

```

## Getting to know you all

For the remainder of the assignment, I need to know three things.

1. What type of data will you be working with for your thesis or research project? For example: matrices of phylogenetic characters, spreadsheets of ecological observations, shape files from 3D scanning, etc. If you aren't actively doing research, what kinds of scientific questions do you find interesting?
= I am working with matrices of phylogenetic characters consisting of molecular data, morphological character information and fossil ages.

2. What are you hoping to get out of class?
= I am hoping to be more fluent in using R for my research works. 

3. Are there topics that aren't on the schedule that you'd be interested to see? 
= No. 
