---
title: "Class_Nov3"
author: "Basanta Khakurel"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Camera Trap Data Analysis and NMDS

## Camera Trap Data Analysis 

Look at some National Camera Traps to identify different organisms occuring in those regions. 
Few different habitats in Florida

Install `mapview` package
```{r}
#remotes::install_github("r-spatial/mapview")
```

Load required packages

```{r}
library(tidyverse)
library(googlesheets4)
library(iNEXT)
library(fields)
library(sf)
library(mapview)
```

Download the data
```{r}
gs4_deauth() #we're pulling some data that are public. So not using password.
cam <- read_sheet("https://docs.google.com/spreadsheets/d/1esrbC-QqoufjlDHXtIYtDcsXovC_x6o6Z4JfBPKvg3s/edit?usp=sharing")
```

Lets look at the data
```{r}
dim(cam)
names(cam)
cam
```

Cleaning up the names 
```{r}
names(cam) <- make.names(names(cam), unique = TRUE) #create good names without spaces

#We need to clean up names and get rid of spaces. These search for spaces in the given column names and replace them with periods
cam$Species.Name <- str_replace_all(cam$Species.Name, " ", ".")
cam$Common.Name <- str_replace_all(cam$Common.Name, " ", ".")
cam$Subproject <- str_replace_all(cam$Subproject, " ", ".")
cam$Subproject <- str_replace_all(cam$Subproject, "/", ".") #there is one subproject name that has a /. Let's just change it to a period for consistency's sake.
```

Summarizing counts and organizing them into subprojects or Forest Types.
Sites in Columns and Species in Rows.
Gathering and spreading the data. 
Collapsing the multiple variable data to one variable statistic.
```{r}
speciesCount <- cam %>% 
  select(Common.Name, Count, Subproject) %>% 
  group_by(Common.Name, Subproject) %>% #Here I group the data by species and the subproject
  summarize(totalCount = sum(Count)) %>% #This now gives the sum of the count of each species per subproject
  spread(Subproject, totalCount) #Now I need to change the orientation so that species are rows and subprojects are columns. This pivots the table to do that. More explanation here: https://rpubs.com/bradleyboehmke/data_wrangling
speciesCount
```

### Abundance for each subproject. 

The number of each individual species in a particular site. 

```{r}
abundance <- cam %>% 
  group_by(Subproject) %>% 
  filter(!is.na(Species.Name)) %>%  #filter out any NAs in the species name, just in case
  summarize(SpeciesAbundance = length(unique(Species.Name)))
abundance
```
This metric has some shortcomings too. 
Could be one species is really dominating with just a minority of other taxa in there. 
Doesn't tell us how much overlap is there between the sites. 


## Species Richness
How much of different species are there in any particular region/sub-project. 

Transform the data to match nicely with iNEXT 
```{r}
speciesCount[is.na(speciesCount)]=0 #This turns the NAs that come from summing 0's into real 0's for analysis
camDiv <- as.data.frame(speciesCount) #Since iNext wants a data frame, turn the tibble into one
rownames(camDiv) <- camDiv$Common.Name #iNext also wants the species ID as row names. 
camDiv <- camDiv[2:6] #Now we can get rid of the species name since it's already in our row name.
```


Calculate Species Richness and Plot it. 
```{r}
richness <- iNEXT(camDiv, q=0, datatype = "abundance") #This calculates hill numbers for the number of observations we have. q changes the expected relationship. 0 will give Richness
ggiNEXT(richness) + theme(legend.position = "right") + theme_classic()
```
$\sum_i..j [P_i log(P_i)]
### Calculate Shannon's Diversity Index
```{r}
diversity_Shan <- iNEXT(camDiv, q=1, datatype = "abundance") #This calculates hill numbers for the number of observations we have. q changes the expected relationship. q=1 gives the Shannon Diversity index. 
ggiNEXT(diversity_Shan)+
  theme(legend.position = "right")+
  theme_classic()
```
Simpsons goes from 0 to 1 
### Calculate Simpson's Diversity Index 
```{r}
diversity_Simps <- iNEXT(camDiv, q=2, datatype = "abundance") #This calculates hill numbers for the number of observations we have. q changes the expected relationship. q=2 gives the Simpsons Diversity index.  
ggiNEXT(diversity_Simps)+
  theme(legend.position = "right")+
  theme_classic()
```

Looking at each diversity calculation within each site 
```{r}
allDiversity <- iNEXT(camDiv, q=c(0, 1, 2), datatype="abundance", endpoint=7000)
```

```{r}
ggiNEXT(allDiversity, type=1, facet.var="Assemblage")+
  theme_classic()
```



Plotting by deployment location and total number of species 
```{r}
ggplot(cam, aes(x=Actual.Lon, y=Actual.Lat, color = Subproject))+
  geom_point()+
  theme_bw()
```

Location of each camera scaled by numbers of animals and colored by the habitat 
```{r}
deployCount <- cam %>% dplyr::select(Common.Name, Count, Deployment.ID, Actual.Lon, Actual.Lat, Subproject) %>% 
  group_by(Deployment.ID, Actual.Lon, Actual.Lat, Subproject) %>% 
  summarize(totalCount = sum(Count))

ggplot(deployCount)+
  geom_point(aes(x=Actual.Lon, y=Actual.Lat, size=totalCount, color=Subproject), alpha=0.5)+
  theme_bw()
```

Making a Zoomable and Interactive map.
```{r}
library(sf)
library(mapview)

#Transform these objects into something that can be plotted with the package {sf}
deployCount_sf <- deployCount %>% st_as_sf(coords = c("Actual.Lon", "Actual.Lat"), crs=4326)

camMap <- mapview(deployCount_sf, zcol="Subproject", cex="totalCount", layer.name = "Habitat Type")
camMap
```

***
***

# NMDS

For each size what we have list of species and the number of observations.  
```{r}
library(vegan)
```


Vegan will expect the species names to be the columns and the site to be the row:

```{r}
speciesCount <- cam %>% 
  select(Common.Name, Count, Subproject) %>% 
  group_by(Common.Name, Subproject) %>% #Here I group the data by species and the subproject
  summarize(totalCount = sum(Count)) %>% #This now gives the sum of the count of each species per subproject
  spread(Common.Name, totalCount, fill = 0) 

speciesMatrix <- data.matrix(speciesCount, rownames.force = unique(cam$Subproject))
```

```{r}
sm <- speciesMatrix[,-1]
```
Perform NMDS
```{r}
cam_trap_NMDS <- metaMDS(sm)
```

Create a stressplot.
```{r}
stressplot(cam_trap_NMDS)
```

```{r}
ordiplot(cam_trap_NMDS,type="n")
orditorp(cam_trap_NMDS,display="species",col="red",air=0.01)
orditorp(cam_trap_NMDS,display="sites",cex=1.25,air=0.01)
```

```{r}
cols <- c('red', 'black', 'blue', 'green', 'orange')
groups <- seq(length(speciesCount$Subproject))

ordiplot(cam_trap_NMDS,type="text")
orditorp(cam_trap_NMDS,display="sites",cex=1.25,air=0.01)
text(cam_trap_NMDS, display = 'sites')
```

To create a more beautiful NMDS plot
```{r}
scores(cam_trap_NMDS) %>%
  cbind(speciesCount) %>% 
  ggplot(aes(x = species.NMDS1, y = species.NMDS2)) + geom_point(aes(color = Subproject)) + stat_ellipse(type= "norm", linetype = 2, geom = "polygon", aes(group = Subproject, fill = Subproject), alpha = 0.3) + annotate("text", x=-2, y=1, label = paste0("stress", format(cam_trap_NMDS$stress, digits=4), hjust =0)) + theme_bw()
```


#### Your Turn 

```{r}
nop <- cam <- read_sheet("https://docs.google.com/spreadsheets/d/18yhZcaiYt6llg84eZYrNEewuLP3xaEyxYQQN__RQ3_4/edit?usp=sharing")
```
