---
title: "Class_Oct3"
author: "Basanta Khakurel"
date: "`r Sys.Date()`"
output: html_document
---

## Class Notes Oct 4, 2022

# Version Control Continued + Linear Modeling

Using github.....

***

# Linear Modeling 
One of the most flexible analysis to do in R
- Using crabs.csv data set

If we know the correlation of one variable with the other, could we predict one variable given that we have the other variable.

$Y = \beta X + \beta_0 + \in$

Linear model formula 

Y is the response variable and X is the predictor variable. B is the relationship coefficient between the predictor and response. 

```{r}
setwd("~/projects/Biological_Data_Anaysis_2022/")
library(tidyverse)
```

```{r}
crabs <- read_csv("data/crabs.csv")
glimpse(crabs)
```

Look at the data using a plot.

```{r}
ggplot(crabs, aes(x = carapace_length, y = body_depth)) + geom_point()
```

create a model_fit variable 
```{r}
model_fit <- lm(body_depth ~ carapace_length, data=crabs)
summary(model_fit)
```

create a plot with a linear model fitting
```{r}
ggplot(crabs, aes(x = carapace_length, y = body_depth)) + geom_point() + geom_smooth(method = "lm", color = "navy", size = 0.5, fill = "deeppink") + labs(x= "Carapace Length(mm)", y="Body Depth(mm)", title = "Carapace Length vs Body Depth") + annotate("text", x =17, y=24, label = "R^2 == 0.966", parse=T, color = "firebrick", size = 5)+theme_bw()
```

Confidence Interval - if our data meets the assumptions of our model, then there is a 95% chance that the true data might be along that interval 


