---
title: "Class_Oct3"
author: "Basanta Khakurel"
date: "`r Sys.Date()`"
output: html_document
---

## Class Notes Oct 4, 2022

# Version Control Continued + Linear Modeling

Using github.....

***

# Linear Modeling 
One of the most flexible analysis to do in R
- Using crabs.csv data set

If we know the correlation of one variable with the other, could we predict one variable given that we have the other variable.

$Y = \beta X + \beta_0 + \in$

Linear model formula 

Y is the response variable and X is the predictor variable. B is the relationship coefficient between the predictor and response. 

```{r}
#setwd("c:/Users/basan/Desktop/Fall 2022/Biological_Data_Anaysis_2022/")
library(tidyverse, broom)
```

```{r eval = TRUE}
crabs <- read_csv("../data/crabs.csv")
glimpse(crabs)
```

Look at the data using a plot.

```{r}
ggplot(crabs, aes(x = carapace_length, y = body_depth)) + geom_point()
```

create a model_fit variable 
```{r}
model_fit <- lm(body_depth ~ carapace_length, data=crabs)
summary(model_fit)
```

create a plot with a linear model fitting
```{r}
ggplot(crabs, aes(x = carapace_length, y = body_depth)) + geom_point() + geom_smooth(method = "lm", color = "navy", size = 0.5, fill = "deeppink") + labs(x= "Carapace Length(mm)", y="Body Depth(mm)", title = "Carapace Length vs Body Depth") + annotate("text", x =17, y=24, label = "R^2 == 0.966", parse=T, color = "firebrick", size = 5)+theme_bw()
```

Confidence Interval - if our data meets the assumptions of our model, then there is a 95% chance that the true data might be along that interval 

---
---
---
## Oct 6 Notes 
```{r}
model_fit_2 <- lm(frontal_lobe ~ body_depth, data = crabs)

summary(model_fit_2)
```

Plot and see the linear model fit 
```{r}
ggplot(crabs, aes(x = Latitude, y = body_depth))+
  geom_point(size = 2)+ 
  geom_smooth(method = "lm",color = "navy", size = 0.5, fill = "deeppink4") 
```


## Multiple Variables 
### Additive Effect

This one looks at a categorical and a numeric additive effect

```{r, eval=FALSE}
model_fit_multi <- lm(body_depth ~ color+carapace_length, data=crabs)

summary(model_fit_multi)

augmented_multi <- augment(model_fit_multi)

#quantiles of quantile plot
qqnorm(augmented_multi.resid)
qqline(augmented_multi$.resid, col="red")

ggplot(crabs, aes(x=carapace_length, y=body_depth, col=color))+
  geom_point()+ 
  scale_color_manual(values = c("blue", "orange"))+ 
  geom_smooth(method = "lm", fill = "green")+ 
  annotate("text", x = 15, y = 25, label = "R^2 == 0.988", parse= T, size = 4, color = "firebrick")+
  theme_bw() 
```

Here we are looking at the plot of carapace width and body depth. 
```{r}
ggplot(crabs, aes(x = carapace_width, y = body_depth)) + geom_point()
```

This one looks at the additive effect of two numerical variables.
```{r, eval=FALSE}
model_fit <- lm(body_depth ~ carapace_width + carapace_length, data = crabs)

summary(model_fit)

augment(model_fit) -> augmented_fit
qqnorm(augmented_fit$.resid)
qqline(augmented_fit$.resid, col="red")

ggplot(crabs, aes(x = carapace_length, y = body_depth, color = carapace_width)) + 
  geom_point(size = 2) +
  labs(x = "Carapace length (mm)", 
       y = "Body depth (mm)",
       color = "Carapace width (mm)") +
  scale_color_distiller(palette = "Greens") +
  annotate("text", x = 20, y = 30, label = "R^2 == 0.977", parse=T, size=5) + 
  theme(legend.position = "bottom")
```


### Interaction Effect

Numeric and Categorical Predictors
```{r, eval=FALSE}
model_fit <- lm(body_depth ~ carapace_length*color, data =crabs)

summary(model_fit)

augmented_fit <- augment(model_fit)
qqnorm(augmented_fit$.resid)
qqline(augmented_fit$.resid, col = "blue")
```

Numeric and Numeric Predictors          

```{r, eval=FALSE}
model_fit <- lm(body_depth ~ carapace_length*carapace_width, data =crabs)

summary(model_fit)

augmented_fit <- augment(model_fit)
qqnorm(augmented_fit$.resid)
qqline(augmented_fit$.resid, col = "blue")

#Create an Interaction Plot
crabs %>% 
  group_by(sex, color) %>% 
  summarize(mean_body_depth = mean(body_depth)) %>% 
ggplot(aes(sex, mean_body_depth, color=color)) +
  geom_point()+
  labs(x = "Sex", y = "Mean Body Depth (mm)", color = "Crab Color") +
  scale_color_manual(values = c("blue", "orange"))+
  geom_path(aes(group=color))+
  annotate("text", x=0.5, y = 15,label = "R^2 == 0.9775", parse = T)
```

```{r, eval=FALSE}
model_fit <- lm(body_depth ~ sex*color, data =crabs)
summary(model_fit)
augmented_fit <- augment(model_fit)
qqnorm(augmented_fit$.resid)
qqline(augmented_fit$.resid, col = "blue")
```

