---
title: "Class_Oct27"
author: "Basanta Khakurel"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

## Class Notes Oct 27, 2022

# Mapping Exploration and Simulations

This week we are going to learn about mapping using the data from GBIF (Global Biodiversity Information Facility). It is the occurrence data we would be plotting in the maps. 

```{r, eval=FALSE}
install.packages("rgbif")
install.packages("leaflet") #custom plotting of data in various maps 
```

```{r}
library(tidyverse)
library(rgbif)
library(leaflet)
library(rotl)
```

This will look for the species and give a unique taxon key.
```{r}
name_suggest("Pituophis melanoleucus")
```

Use the returned taxon key to get some occurence data for the species. 
```{r}
occurences <- occ_search(taxonKey = 9187223, limit = 50)
```

remove any Na values from the dataset. 
```{r}
no_na <- occurences$data %>%
  select(scientificName, decimalLatitude, decimalLongitude) %>% 
  drop_na()
```

Plot the occurrence data into a map. The popus are the scientific names.  
```{r}
k <- leaflet::leaflet(no_na) %>% 
  addTiles() %>% 
  addMarkers(~decimalLongitude, ~decimalLatitude, popup = no_na$scientificName)
```

---

Multiple Species in the plot 
```{r}
ants <- c("Martialis", "Atta", "Ectatomma", "Xymmer", "Aneuretus")
```

This loop looks at each of the ant genera and look up in the database.  
Getting a vector of data frames. 

```{r}
search_results <- list() #create an empty vector
for (ant in ants){
  query <- name_suggest(ant, rank = "genus")
  search <- occ_search(taxonKey = query$data$key, limit = 10)
  search_results[[ant]] <- search$data
  print(search)
  }
```

Add all the data frames together. 

```{r}
mega_df <- bind_rows(search_results)
```

```{r}
no_na <- mega_df %>%
  select(scientificName, decimalLatitude, decimalLongitude) %>% 
  drop_na()
```

```{r}
leaflet(no_na) %>% 
  addTiles() %>% 
  addMarkers(~decimalLongitude, ~decimalLatitude, popup = no_na$scientificName)
```

---

## Choosing your own adventure (November 1)

Circles instead of Tiles 
```{r}
leaflet(no_na) %>% 
  addTiles() %>% 
  addCircles(~decimalLongitude, ~decimalLatitude)
```

Draw Point sizes from a distribution 
```{r}
leaflet(no_na) %>% 
  addTiles() %>% 
  addCircleMarkers(~decimalLongitude, ~decimalLatitude, radius = runif(100, 4, 10), color = c('red'))
```

Cluster Observations
```{r}
leaflet(no_na) %>%
  addTiles() %>% 
  addMarkers(~decimalLongitude, ~decimalLatitude, clusterOptions = markerClusterOptions())
```

Size points by number of observations
```{r}
no_na %>% 
  group_by(decimalLatitude) %>% 
  mutate(n_areas = n())  %>% 
  leaflet() %>%
  addTiles() %>% 
  addCircleMarkers(~decimalLongitude, ~decimalLatitude, radius = ~n_areas, color = c('red'))
```

Color Points by Species 
```{r}
pal <- colorBin(
  palette = "Blues",
  no_na$scientificName,
  pretty = TRUE)

levs <- factor(no_na$scientificName)
factpal <- colorFactor(topo.colors(5), levs)

no_na %>% 
  group_by(scientificName) %>% 
  leaflet() %>%
  addTiles() %>% 
  addCircleMarkers(
    ~decimalLongitude, 
    ~decimalLatitude,
    color = ~factpal(scientificName),
    stroke = FALSE, fillOpacity = 0.5
  ) 
```

Set view Width 
```{r}
pal <- colorBin(
  palette = "Blues",
  no_na$scientificName,
  pretty = TRUE)

levs <- factor(no_na$scientificName)
factpal <- colorFactor(topo.colors(5), levs)

no_na %>% 
  group_by(scientificName) %>% 
  leaflet() %>%
  addTiles() %>% 
  addCircleMarkers(
    ~decimalLongitude, 
    ~decimalLatitude,
    color = ~factpal(scientificName),
    stroke = FALSE, fillOpacity = 0.5
  ) %>% 
    setView( lng = -100,
            lat = 20,
            zoom = 11 ) %>%
  setMaxBounds( lng1 = -100,
                lat1 = 19.432241,
                lng2 = -98,
                lat2 = 20 )
```

***
This is after show and tell. 


## Mapping a tree to space 
```
devtools::install_github("liamrevell/phytools")
library(phytools)
```

```{R}
split_names <- no_na %>% 
mutate(genus = map_chr(scientificName, function(s) strsplit(s, " ")[[1]][1]))%>% mutate(species = map_chr(scientificName, function(s) strsplit(s, " ")[[1]][2])) %>% 
unite(col = genusSpecies, genus, species)
```


```{r}
# Use ROTL to resolve names

no_bold <- split_names[ grep("BOLD", split_names$genusSpecies, invert = TRUE) , ]
```

```{r}
a_couple_ants <- c("Martialis_heureka", "Ectatomma_ruidum", "Ectatomma_tuberculatum", "Aneuretus_simoni", "Ectatomma_opaciventre")
```

```{r, echo = FALSE}
subset_data <- no_bold %>% 
    filter(genusSpecies %in% a_couple_ants)
```

```{r}
reconciled_names <- rotl::tnrs_match_names(unique(subset_data$genusSpecies))
good_names <-  reconciled_names %>%
  drop_na()
  
tree <- rotl::tol_induced_subtree(good_names$ott_id, label="name")

plot(tree)
```

```{r,eval=FALSE}
only_lat_long <- subset_data %>% 
  select(decimalLatitude, decimalLongitude) %>% 
  as.matrix()

rownames(only_lat_long) <- subset_data$genusSpecies
```

```{r}
library(viridis)
```

```{r, eval=FALSE}
color_selection <- setNames(sample(viridis(n=Ntip(tree))),
    tree$tip.label)

obj<-phylo.to.map(tree,only_lat_long, plot=FALSE, direction="rightwards")
plot(obj)
```
This does not have any branch lenghts. 

Lets try again with some changes. 
```{r}
library(ape)
```

```{r}
tree <- multi2di(tree)
```

```{r}
tree$edge.length <- rexp(tree$Nnode*2)
tree$edge.length 
```

```{r,eval=FALSE}
plot_object <- phylo.to.map(tree, only_lat_long, plot = FALSE)
plot_object
plot(plot_object, direction = "rightwards", colors = color_selection, cex.points=c(0.1), lwd=c(3,1),ftype="off")
```

Simulation Discussion

