---
title: "Class_Week of Sept 20"
author: "Basanta Khakurel"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Class Notes Sept 21, 2022

# Functions 
## definition, body and return value

```{r}
calc_vol <- function(len = 1, width = 1, height = 1){
  volume <- len * width * height
  return(volume)
}
```

```{r, eval = FALSE}
calc_vol(45, 23, 45)
```
Local scope is where the variables used in the function exists, it doesn't get saved to the Global Environment to save memory 


```{r}
calc_mass_index <- function(weight =1, kg = FALSE, foot_length =1){
  if (kg == FALSE){
    weight <- weight*.454
  }
  mass_index <- weight/foot_length
  return(mass_index)
}
```

```{r}
calc_mass_index <- function(weight =1, foot_length =1){
  mass_index <- (weight*.454)/foot_length
  return(mass_index)
}
```


```
calc_density <- function(mass = NA, unit_m = c("kg", "lbs"), volume = NA, unit_v = c("cm3, m3, in3, mm3")){
  if (unit_m == "lbs"){
    mass <- mass*0.454
  }
  if (unit_v == "cm3"){
    volume <- volume/1000000
  } else if (unit_v == "in3"){
      volume <- volume/61020
  } else if (unit_V == "mm3"){
      volume <- volume/1000000000
    }
  density <- mass/volume
  return(density)
}
```


```{r}
calc_mass <- function(vol){
  mass <- 2.65 * vol^0.9
  return(mass)
}
```

```{r}
cal_den <- function(len =1, widt=1, height=1){
  vol <- calc_vol(len, widt, height)
  mass <- calc_mass(vol)
  density <- mass/vol
  return(density)
}
```
---
### Testing or Code Reliability

*What is a good number?
+ bounding (positive/negative, decimals)
+ Integer or Double
+ Precision

*What is a good density?
+ density < volume
+ is relative to mass

#### This will test the accuracy/reliability of the function
```{r}
cal_den <- function(len =1, widt=1, height=1){
  vol <- calc_vol(len, widt, height)
  mass <- calc_mass(vol)
  density <- mass/vol
  if (density > 0){
    return(density)
  } else {
    print("Density impossibly small! Check your inputs.")
  }
}
```



```{r}
#this code checks for math errors
cal_den <- function(len =1, widt=1, height=1){
  vol <- calc_vol(len, widt, height)
  mass <- calc_mass(vol)
  density <- mass*vol
  if (density < vol){
    return(density)
  } else {
    print("Density is larger than volume! Check your inputs.")
  }
}
```


`data_cleaning` function: reads a data frame, some data cleaning, return the cleaned data 
```{r}
data_cleaning <- function(path_to_data){
  data <- read_csv(path_to_data)
  #return(View(data))
  data_clean <- na.omit(data)
  if (sum(is.na(data_clean)) == 0) {
    return(data_clean)
  } else {
    print("NAs still present!!")
  }
}
```

```{r}
#data_cleaning("../data/portal_data_joined.csv")
```
What is a good way to test that it worked?

```anyNA```


https://docs.ropensci.org/treedata.table/
Look at Development, Maintainance and Peer Review